---
title: Exploratory Data Analysis
format: typst
---

```{r}
library(tidyverse)
library(readr)
library(sf)
library(tigris)
library(dplyr)
library(ggplot2)
library(stats)
precinct_2024_general_election <- read.csv("data/g24_sov_by_g24_svprec.csv")

simpler_precinct <- precinct_2024_general_election |>
    select("COUNTY", "TOTVOTE", "CDDIST", "ASSDEM01", "ASSDEM02", "ASSREP01",
    "ASSREP02", "CNGDEM01", "CNGDEM02", "CNGREP01", "CNGREP02",
    "PRSDEM01",  "PRSREP01", "SENDEM01", "SENDEM02",
    "SENREP01", "SENREP02", "USSDEM01", "USSREP01") |>
        filter(
            !if_any(
                .cols = -COUNTY, 
                .fns = ~ . == "***"
                )) |>
                    mutate(
                        across(
                            .cols = -COUNTY, 
                            .fns = ~ as.integer(as.character(.))
                        )
                    )


```

### Question 1

How many votes for the presidential democrat vs republican candidate had a less than 1 percent difference in vote count between them?

### Answer 1

```{r}
closest_race_analysis <- simpler_precinct |>

  mutate(
    Contest_Total = PRSDEM01 + PRSREP01,
    Raw_Difference = abs(PRSDEM01 - PRSREP01),
    Pct_Margin = (Raw_Difference / Contest_Total) * 100
  )

closest_precinct <- closest_race_analysis |>
  arrange(Pct_Margin)
```

```{r}
count_under_1_percent <- closest_race_analysis |>
  filter(Pct_Margin < 1) |>
  count()
count_under_1_percent
```

782


### Question 2

What is the range of total sizes of the precincts? 

### Answer 2

```{r}
precinct_size_summary <- simpler_precinct |>
  summarise(
    Min_Total_Votes = min(TOTVOTE),
    Max_Total_Votes = max(TOTVOTE),
    Mean_Total_Votes = mean(TOTVOTE),
    Median_Total_Votes = median(TOTVOTE),
    Range_of_Votes = Max_Total_Votes - Min_Total_Votes
  )
print(precinct_size_summary)
```

The range goes from a vote count of 0 to 3793980, with a mean of 2424.274 votes and median of 121 votes.

### Question 3

How large were the largest counties and how small were the smallest?

### Answer 3


```{r}
county_precinct_counts <- simpler_precinct |>
  group_by(COUNTY) |>
  count(name = "Precinct_Count") |>
  arrange(desc(Precinct_Count))
print(county_precinct_counts)
```

Largest county is labeled county 19 with 5937 precincts.


```{r}
county_precinct_counts <- simpler_precinct |>
  group_by(COUNTY) |>
  count(name = "Precinct_Count") |>
  arrange((Precinct_Count))
print(county_precinct_counts)
```

Smallest is county 2 with 19 precincts.