---
title: "Practice Worksheet: Gerrymandering"
format: typst
---

Recall the 3 quantitative approaches to measuring gerrymandering discussed in lecture:\
1. Mean - Median Score (Grofman, 1983)\
2. Efficiency Gap (Stephanopolous and McGee, 2015)\
3. Ensembles / Extreme Outliers\
Use the following table to practice applying the first two techniques.\

| District | D Votes | R Votes | Result  |
|:---------:|:-------:|:-------:|:--------|
| 1 | 62 | 38 | D wins |
| 2 | 55 | 45 | D wins |
| 3 | 44 | 56 | R wins |
| 4 | 47 | 53 | R wins |
| 5 | 46 | 54 | R wins |
| **Total** | **254** | **246** |          |

```{r}
#| message: false
#| echo: false
library(tidyverse)
votes <- tibble(District = 1:5,
                d_votes = c(62, 55, 44, 47, 46),
                r_votes = c(38, 45, 56, 53, 54),
                result = c(rep("D wins", 2), rep("R wins", 3)))
```

### Q1. Mean-Median Score

Calculate the mean-median score associated with this particular district breakdown of votes.

```{r}
# add code for Q1 here
mean_votes_dem <- mean(votes$d_votes)
median_votes_dem <- median(votes$d_votes)
mean_minus_median_dem <- mean_votes_dem - median_votes_dem


mean_votes_rep <- mean(votes$r_votes)
median_votes_rep <- median(votes$r_votes)
mean_minus_median_rep <- mean_votes_rep - median_votes_rep

mean_minus_median_dem

```

### Q2. Efficiency Gap

Calculate the Efficiency Gap (EG) associated with this particular district breakdown of votes.

*Tip*: it may be helpful to write a function that takes a vector of votes across districts for party A and vector of votes for the party B and returns a vector of the number votes wasted by party A across districts.


```{r}
# add code for Q2 here

# wasted_votes <- function(x) {
#     num_wasted <- 0
#     if (x >= 51) {
#         num_wasted <- (x - 51) 
#     } else {
#         num_wasted <- x
#     }
#     print(num_wasted)
# }

# votes_eg <- votes |>
#     mutate(d_wasted = sapply())


```